---
title: "Rename"
output: html_notebook
---

# Remove date range specification from Volume-Occupancy file names

Originally, start date and end date were specified as a part of file name within corridor_pull_par function.  This was determined to be a nuisance for a variety of reasons; thus date range specification is removed here from all file names.

```{r}
folder <- "N:/CommDev/Research/Public/MTS/Metro Loop Detectors/Volume-Occupancy Data/2018/I-494"

hwys <- list.files(folder, full.names = T)

map(hwys, function(eachPath){ 
   file.rename(from = eachPath, to = sub(pattern = "20180101-20181231", paste0(""), eachPath))
 })

```

# Remove year attribution

```{r}

folder <- "N:/CommDev/Research/Public/MTS/Metro Loop Detectors/Volume-Occupancy Data/2018/T.H.252"

hwys <- list.files(folder, full.names = T)

map(hwys, function(eachPath){ 
   file.rename(from = eachPath, to = sub(pattern = "2018 ", paste0(""), eachPath))
 })

```

# Get rid of space between extension and .csv in file name

Next, year is added as prefix (in case files are moved, the year is a helpful differentiator).

```{r}
folder <- "N:\\CommDev\\Research\\Public\\MTS\\Metro Loop Detectors\\Volume-Occupancy Data\\2018\\T.H.610"

sensors <- list.files(folder, full.names = T)

map(sensors, function(eachPath){ 
   file.rename(from = eachPath, to = sub(pattern = "30 .csv", paste0("30.csv"), eachPath))
 })

```

# Get rid of space between extension and .csv in file name

Next, year is added as prefix (in case files are moved, the year is a helpful differentiator).

```{r}
folder <- "N:\\CommDev\\Research\\Public\\MTS\\Metro Loop Detectors\\Volume-Occupancy Data\\2018\\U.S.169"

sensors <- list.files(folder, full.names = T)

map(sensors, function(eachPath){ 
   file.rename(from = eachPath, to = sub(pattern = ".csv", paste0(" c30.csv"), eachPath))
 })

```
