---
title: "XML Configuration Conversion and Calculating Distances Between Nodes"
output: html_document
---

## Conversion, excluding empty rnodes (no detectors)

Code chunk below is used for creating configuration csv file from xml.

xml is obtained from:  http://www.dot.state.mn.us/tmc/trafficinfo/developers.html

Click on "metro_config.xml.gz" link and unzip.

```{r}
library(stringr)
library(tidyverse)
library(xml2)
library(data.table)
library(geosphere)

# temp <- tempfile()
# download.file("http://data.dot.state.mn.us/iris_xml/metro_config.xml.gz", temp) # Download zipped gz file
# data <- read.table(unz(temp, "metro_config.xml"))
# unlink(temp)

metro_config <- read_xml('Configuration Data/xml files/Configuration 2019-07-08.xml')
metro_config_detectors <- data.frame(xml_path(xml_find_all(metro_config, "//detector")))

#Grab all detector paths
detector_rnode_paths <- metro_config_detectors %>%
  mutate(Detector_Path = metro_config_detectors[[1]]) %>%
  separate(1, into = c('front', 'tms_config', 'device', 'rnode', 'Detector'), sep = '/') %>%
  unite(Rnode_path, front, tms_config, device, rnode, sep = '/') %>%
  mutate(Rnode_path = trimws(Rnode_path)) %>%
  mutate(Corridor_path = Rnode_path) %>%
  separate(Corridor_path, into = c('front', 'tms_config', 'device', 'rnode'), sep = '/') %>%
  unite(Corridor_path, front, tms_config, device, sep = '/') %>%
  mutate(Corridor_path = trimws(Corridor_path))

#Grab all detector attributes and combine with paths
detectors <- xml_find_all(metro_config, '//detector')
detector_names <- trimws(xml_attr(detectors, 'name'))
detector_labels <- trimws(xml_attr(detectors, 'label'))
detector_categories <- trimws(xml_attr(detectors, 'category'))
detector_lanes <- trimws(xml_attr(detectors, 'lane'))
detector_fields <- trimws(xml_attr(detectors, 'field'))
detector_abandoned <- trimws(xml_attr(detectors, 'abandoned'))

detector_attr <- data.frame(Detector_name = detector_names, Detector_label = detector_labels, Detector_category = detector_categories, Detector_lane = detector_lanes, Detector_field = detector_fields, Detector_abandoned = detector_abandoned)

detector_rnode_paths_attrs <- bind_cols(detector_rnode_paths, detector_attr)

#Grab all r_node paths
metro_config_rnodes <- data.frame(xml_path(xml_find_all(metro_config, "//r_node")))

rnode_paths <- metro_config_rnodes %>%
  mutate(Rnode_path = metro_config_rnodes[[1]]) %>%
  select(Rnode_path)

#Grab all r_node attributes and combine with paths
rnodes <- xml_find_all(metro_config, '//r_node')
rnode_names <- trimws(xml_attr(rnodes, 'name'))
rnode_n_types <- trimws(xml_attr(rnodes, 'n_type'))
rnode_transitions <- trimws(xml_attr(rnodes, 'transition'))
rnode_labels <- trimws(xml_attr(rnodes, 'label'))
rnode_lons <- trimws(xml_attr(rnodes, 'lon'))
rnode_lats <- trimws(xml_attr(rnodes, 'lat'))
rnode_lanes <- trimws(xml_attr(rnodes, 'lanes'))
rnode_shifts <- trimws(xml_attr(rnodes, 'shift'))
rnode_s_limits <- trimws(xml_attr(rnodes, 's_limit'))
rnode_stations <- trimws(xml_attr(rnodes, 'station_id'))
rnode_attach_sides <- trimws(xml_attr(rnodes, 'attach_side'))

rnode_attr <- data.frame(Rnode_Name = rnode_names, Rnode_n_type = rnode_n_types, Rnode_transition = rnode_transitions, Rnode_label = rnode_labels, Rnode_lon = rnode_lons, Rnode_lat = rnode_lats, Rnode_lanes = rnode_lanes, Rnode_shift = rnode_shifts, Rnode_slimit = rnode_s_limits, Rnode_station_id = rnode_stations, Rnode_attach_side = rnode_attach_sides)

rnode_paths_attrs <- bind_cols(rnode_paths, rnode_attr)

detector_rnodes_full <- left_join(detector_rnode_paths_attrs, rnode_paths_attrs, by = c('Rnode_path'))

#Grab all corridors
metro_config_corridors <- data.frame(xml_path(xml_find_all(metro_config, "//corridor")))

corridor_paths <- metro_config_corridors %>%
  mutate(Corridor_path = metro_config_corridors[[1]]) %>%
  select(Corridor_path)

#Grab corridor path attributes
corridors <- xml_find_all(metro_config, '//corridor')
corridor_routes <- trimws(xml_attr(corridors, 'route'))
corridor_dirs <- trimws(xml_attr(corridors, 'dir'))

corr_attrs <- data.frame(Corridor_route = corridor_routes, Corridor_dir = corridor_dirs)

corr_paths_attrs <- bind_cols(corridor_paths, corr_attrs)

detector_rnode_corrs <- left_join(detector_rnodes_full, corr_paths_attrs, by = c('Corridor_path'))

detector_rnode_corrs_tidy <- detector_rnode_corrs %>%
  select(-Rnode_path, -Detector, -Detector_Path, -Corridor_path, -rnode)

```

## Conversion, including empty rnodes

This configuration is used for visualization purposes, as empty rnodes better connect detectors.

```{r}
library(stringr)
library(tidyverse)
library(xml2)
metro_config <- read_xml('metro_config.xml')
metro_config_detectors <- data.frame(xml_path(xml_find_all(metro_config, "//detector")))

#Grab all detector paths
detector_rnode_paths <- metro_config_detectors %>%
  mutate(Detector_Path = metro_config_detectors[[1]]) %>%
  separate(1, into = c('front', 'tms_config', 'device', 'rnode', 'Detector'), sep = '/') %>%
  unite(Rnode_path, front, tms_config, device, rnode, sep = '/') %>%
  mutate(Rnode_path = trimws(Rnode_path))

#Grab all detector attributes and combine with paths
detectors <- xml_find_all(metro_config, '//detector')
detector_names <- trimws(xml_attr(detectors, 'name'))
detector_labels <- trimws(xml_attr(detectors, 'label'))
detector_categories <- trimws(xml_attr(detectors, 'category'))
detector_lanes <- trimws(xml_attr(detectors, 'lane'))
detector_fields <- trimws(xml_attr(detectors, 'field'))
detector_abandoned <- trimws(xml_attr(detectors, 'abandoned'))

detector_attr <- data.frame(Detector_name = detector_names, Detector_label = detector_labels, Detector_category = detector_categories, Detector_lane = detector_lanes, Detector_field = detector_fields, Detector_abandoned = detector_abandoned)

detector_rnode_paths_attrs <- bind_cols(detector_rnode_paths, detector_attr)

#Grab all r_node paths
metro_config_rnodes <- data.frame(xml_path(xml_find_all(metro_config, "//r_node")))

rnode_paths <- metro_config_rnodes %>%
  mutate(Rnode_path = metro_config_rnodes[[1]]) %>%
  separate(1, into = c('front', 'tms_config', 'device', 'rnode', 'Detector'), sep = '/') %>%
  unite(Corridor_path, front, tms_config, device, sep = '/') %>%
  mutate(Corridor_path = trimws(Corridor_path)) %>%
  select(Corridor_path, Rnode_path)

#Grab all r_node attributes and combine with paths
rnodes <- xml_find_all(metro_config, '//r_node')
rnode_names <- trimws(xml_attr(rnodes, 'name'))
rnode_n_types <- trimws(xml_attr(rnodes, 'n_type'))
rnode_transitions <- trimws(xml_attr(rnodes, 'transition'))
rnode_labels <- trimws(xml_attr(rnodes, 'label'))
rnode_lons <- trimws(xml_attr(rnodes, 'lon'))
rnode_lats <- trimws(xml_attr(rnodes, 'lat'))
rnode_lanes <- trimws(xml_attr(rnodes, 'lanes'))
rnode_shifts <- trimws(xml_attr(rnodes, 'shift'))
rnode_s_limits <- trimws(xml_attr(rnodes, 's_limit'))
rnode_stations <- trimws(xml_attr(rnodes, 'station_id'))
rnode_attach_sides <- trimws(xml_attr(rnodes, 'attach_side'))

rnode_attr <- data.frame(Rnode_Name = rnode_names, Rnode_n_type = rnode_n_types, Rnode_transition = rnode_transitions, Rnode_label = rnode_labels, Rnode_lon = rnode_lons, Rnode_lat = rnode_lats, Rnode_lanes = rnode_lanes, Rnode_shift = rnode_shifts, Rnode_slimit = rnode_s_limits, Rnode_station_id = rnode_stations, Rnode_attach_side = rnode_attach_sides)

rnode_paths_attrs <- bind_cols(rnode_paths, rnode_attr)

detector_rnodes_full <- left_join(rnode_paths_attrs, detector_rnode_paths_attrs, by = c('Rnode_path'))

#Grab all corridors
metro_config_corridors <- data.frame(xml_path(xml_find_all(metro_config, "//corridor")))

corridor_paths <- metro_config_corridors %>%
  mutate(Corridor_path = metro_config_corridors[[1]]) %>%
  select(Corridor_path)

#Grab corridor path attributes
corridors <- xml_find_all(metro_config, '//corridor')
corridor_routes <- trimws(xml_attr(corridors, 'route'))
corridor_dirs <- trimws(xml_attr(corridors, 'dir'))

corr_attrs <- data.frame(Corridor_route = corridor_routes, Corridor_dir = corridor_dirs)

corr_paths_attrs <- bind_cols(corridor_paths, corr_attrs)

detector_rnode_corrs <- left_join(detector_rnodes_full, corr_paths_attrs, by = c('Corridor_path'))

detector_rnode_corrs_tidy <- detector_rnode_corrs %>%
  select(-Rnode_path, -Detector, -Detector_Path, -Corridor_path)

#write_csv(detector_rnode_corrs_tidy, 'Configuration of Metro Detectors.csv')
```

## Create a corridor cateogry and corridor index

Can be done for both converted xmls.

```{r}
detector_rnode_corrs_tidy_ii <- detector_rnode_corrs_tidy %>%
  mutate(Rnode_lat = as.character(Rnode_lat),
         Rnode_lon = as.character(Rnode_lon)) %>%
  mutate(Rnode_lat = as.numeric(Rnode_lat),
         Rnode_lon = as.numeric(Rnode_lon)) %>%
  mutate(Corridor_category = ifelse(Corridor_route == 'I-35' & Rnode_lat > 45, 'I35 north of cities',
                                    ifelse(Corridor_route == 'I-35' & Rnode_lat <= 45, 'I35 south of cities',
                                           ifelse(Corridor_route == 'T.H.5' & Rnode_lon < -93.3, '5 west of cities',
                                                  ifelse(Corridor_route == 'T.H.5' & Rnode_lon > -93.3, '5 east of cities', 'Other'))))) %>%
  group_by(Corridor_route, Corridor_dir, Corridor_category) %>%
  mutate(Corridor_index = row_number())

```

## Calculate distance between upstream detector and detector of interest (use converted xml that excludes empty rnodes)

Where upstream detector does not exist, or where distance is beyond 3 miles, interpolate so as not to assume large VMT merely because vehicles crossed an isolated detector.

```{r}
# Select stations (want distance between stations, not between ramps etc.)
config_stations <- detector_rnode_corrs_tidy_ii %>%
  filter(Rnode_n_type == "Station")

# Conflate lanes so they have the same upstream detector; create index
config_conflated <- detector_rnode_corrs_tidy_ii %>%
  filter(Rnode_n_type == "Station") %>%
  select(Corridor_route, Corridor_dir, Corridor_category, Rnode_lat, Rnode_lon) %>%
  unique() %>%
  group_by(Corridor_route, Corridor_dir, Corridor_category) %>%
  mutate(Corridor_conflated_index = row_number())

# Join index to full dataset
config_pre_lag <- left_join(config_stations, config_conflated, by = c("Corridor_route", "Corridor_dir", "Corridor_category", "Rnode_lat", "Rnode_lon"))

# Create lagged index in order to connect lat/longs of upstream detector to detector of interest
config_lag <- config_conflated %>%
  ungroup() %>%
  mutate(Corridor_conflated_index = Corridor_conflated_index - 1) %>%
  rename(Rnode_lat_upstream = Rnode_lat,
         Rnode_lon_upstream = Rnode_lon)

# Join upstream detectors to full detector dataset
config_full <- left_join(config_pre_lag, config_lag, by = c("Corridor_route", "Corridor_conflated_index", "Corridor_dir", "Corridor_category"))

# Find distance between upstream detector and detector of interest for every detector in dataset - for loop vs. purrr::map
# config_length <- c(1:nrow(config_stations))
# dist_results <- list()
# 
# for (i in config_length) {
# 
# config_slice <- config_full %>%
#   ungroup() %>%
#   slice(i)
# 
# config_x <- c(config_slice$Rnode_lon, config_slice$Rnode_lat)
# config_y <- c(config_slice$Rnode_lon_upstream, config_slice$Rnode_lat_upstream)
# 
# dist_results[i] <- distm(config_x, config_y, fun = distHaversine)/1609.344
# }

config_length <- c(1:nrow(config_stations))
dist_results <- list()

dist_results <- purrr::map(config_length, function(i){
  config_slice <- config_full %>%
    ungroup() %>%
    slice(i)

config_x <- c(config_slice$Rnode_lon, config_slice$Rnode_lat)
config_y <- c(config_slice$Rnode_lon_upstream, config_slice$Rnode_lat_upstream)

distm(config_x, config_y, fun = distHaversine)/1609.344
})

options(scipen = 999)
distance_df <- data.frame(matrix(unlist(dist_results), nrow = length(dist_results), byrow = T)) %>%
  rename(distance = 1)

config_dist <- bind_cols(config_stations, distance_df)

# If there is no upstream detector, or distance attributed is greater than 1.5 mi, interpolate with the corridor median
# If, after interpolating with corridor median, distance is still missing (ie NA for entire corridor), or distance attributed is greater than 3 (want a conservative assumption for how many miles of travel volume at a particular sensor depicts), interpolate with metro-wide network median
# As of July 2019, this results in a max distance, metro-wise of 1.45 mi
config_dist_inter <- config_dist %>%
  group_by(Corridor_route, Corridor_dir, Corridor_category) %>%
  mutate(distance = ifelse(is.na(distance) | distance > 1.5, median(distance, na.rm = T), distance)) %>%
  ungroup() %>%
  mutate(distance = ifelse(is.na(distance) | distance > 3, median(distance, na.rm = T), distance))

# Check interpolation interpolated all NAs
config_dist_inter %>%
  filter(is.na(distance))

fwrite(config_dist_inter, "Configuration Data/Configuration of Metro Detectors 2019-07-08.csv")
```

# Create SpatialLinesDataFrame using raster, and sf

```{r}
# library(sp)
# library(raster)
# library(leaflet)
# library(sf)
# library(tidyverse)
# 
# config_df <- config %>%
#   unite(Corridor_ID, Corridor_route, Corridor_dir, Corridor_category) %>%
#   group_by(Corridor_ID) %>%
#   summarize(Detectors = paste(Detector_name, collapse = ", ")) %>%
#   rename(ID = Corridor_ID) %>%
#   as.data.frame()
# 
# config_ln <- config %>%
#   unite(Corridor_ID, Corridor_route, Corridor_dir, Corridor_category) %>%
#   dplyr::select(Rnode_lon, Rnode_lat, Corridor_ID)
# 
# config_split <- split(config_ln, config_ln$Corridor_ID)
# 
# config_line <- vector("list", length(config_split))
# arg_length <- c(1:length(config_split))
# 
# create_line <- function(i) {
#   
#   config_1 <- config_split[[i]] %>% dplyr::select(-Corridor_ID)
#   config_line[[i]] <<- Line(config_1)
# }
# 
# # Create line object
# config_line <- map(arg_length, create_line)
# 
# config_lines <- vector("list", length(config_line))
# 
# # Add ID to Line object
# create_lines_id <- function(i) {
#   
#  config_1 <- config_line[[i]]
#  config_lines[[i]] <- Lines(config_1, ID = config_df %>% slice(i) %>% dplyr::select(ID))
# }
# 
# config_lines <- map(arg_length, create_lines_id)
# 
# config_lines_sp <- SpatialLines(config_lines)
# config_spldf <- SpatialLinesDataFrame(config_lines_sp, config_df, match.ID = "ID")
# 
# config_spldf@proj4string <- CRS("+init=epsg:4267")
# 
# writeOGR(config_spldf, "Configuration Data", "Configuration Test.shp", "ESRI Shapefile")
# 
# plot(config_spldf)
```

# Create an sp lines object

```{r}

# config_df <- config %>%
#   unite(Corridor_ID, Corridor_route, Corridor_dir, Corridor_category) %>%
#   group_by(Corridor_ID) %>%
#   summarize(Detectors = paste(Detector_name, collapse = ", ")) %>%
#   rename(ID = Corridor_ID) %>%
#   as.data.frame()
# 
# config_ln <- config %>%
#   unite(Corridor_ID, Corridor_route, Corridor_dir, Corridor_category) %>%
#   dplyr::select(Rnode_lon, Rnode_lat, Corridor_ID)
# 
# config_split <- split(config_ln, config_ln$Corridor_ID)
# 
# config_line <- vector("list", length(config_split))
# arg_length <- c(1:length(config_split))
# 
# create_line_st <- function(i) {
#   
#   config_1 <- as.matrix(config_split[[i]] %>% dplyr::select(-Corridor_ID))
#   config_line[[i]] <- st_linestring(config_1)
# }
# 
# # Create line object
# config_line <- map(arg_length, create_line_st)
# 
# (ln_strg <- st_sfc(config_line))
# lnstr_sfg1 <- st_linestring(matrix(runif(6), ncol=2)) 
# lnstr_sfg2 <- st_linestring(matrix(runif(6), ncol=2)) 
# (lnstr_sfc <- st_sfc(lnstr_sfg1, lnstr_sfg2))
# (lnstr_sf <- st_sf(dfr , lnstr_sfc))
# (config_sf <- st_sf(config_df, ln_strg))
# 
# config_sf@proj4string <- CRS("+init=epsg:4267")
# st_write(config_sf, "Configuration Data/Config Sf Test.shp")
# 


```

# Create SpatialLinesDataFrame configuration using inlabru

```{r}
library(inlabru)

# Conversion to SpatialLinesDataFrame without CRS
config_sline <- config_full %>%
  filter(!is.na(Rnode_lat_upstream) & !is.na(Rnode_lon_upstream)) # Remove nodes with no upstream detector (no line can be created)

spl <- sline(config_sline, start.cols = c("Rnode_lon", "Rnode_lat"),
             end.cols = c("Rnode_lon_upstream", "Rnode_lat_upstream"))

# Add CRS projection
proj4string(spl) <- CRS("+init=epsg:4267")

writeOGR(spl, "Configuration Data", "Sline test", "ESRI Shapefile")
```


